Tesseroids: forward modeling gravitational fields in spherical coordinates

### 2014-12-08 manuscript.tex
Draft of accuracy section

$D$ controls both accuracy of the integration and the computation time.

\citet{Li2011} provide a few examples of values for this ratio in the case of
the vertical component of gravitational attraction.

@@

>>>>

$D$ controls both the accuracy of the integration and the computation time.

\citet{Li2011} provide a few examples of values for this ratio in the case of
the vertical component of the gravitational attraction.

@@

The computation point can be considered at any point with equal radial
coordinate located outside the shell because of the symmetry of the shell.

We used a spherical shell with a thickness of 1 km.

The density of the shell is $2670\ kg.m^{-3}$.

The bottom of the shell is at height 0 km above the reference sphere.

The top of the shell is at 1 km height.

The largest errors in the GLQ integration occur directly on top of the
tesseroid but not exactly at it's center, as can be seen in
Figure~\ref{fig:glqerrorsample}.

We discretized the spherical shell into tesseroids
with dimensions of $1^\circ \times 1^\circ \times 1\ km$.

We calculate the effect of a tesseroid model of the spherical shell on a
regular grid.

The grid is at a constant height of 2 km (1 km above the shell).

The grid is located on top of a tesseroid located at the North pole.

We repeated the tesseroid calculations with values of the distance-size ratio
varying from 0 (i.e., no divisions) to 10 in 0.5 intervals.

Figure~\ref{fig:tess_vs_shell} shows the maximum difference between the
shell and tesseroid fields as a function of the distance-size ratio.

Differences are given as a percentage of the shell value.

The curves allow us to choose a value of the distance-size ratio that
produces an acceptable error.

We chose 0.1\% as the maximum allowed error.

We performed the computations again for three different scenarios: at 260 km
height (the mean orbit height of the GOCE satellite), on top of a tesseroid at
the equator, and for a $30^\circ \times 30^\circ \times 1\ km$ tesseroid model
of the shell.

We performed these computations to investigate if the GLQ error decays the same
way under different scenarios.

The error curves for the gravitational potential show increased errors at the
equator and for a $30^\circ$ tesseroid.

All errors will be smaller than 0.1\% if one uses a distance-size ratio of 1.

Even in the extreme case of a $30^\circ$ tesseroid with computation points at a
height of 1 km above the tesseroid.

A distance-size ratio of 1.5 is sufficient to achieve errors below 0.1\% in
$g_z$.

A distance-size ratio of 8 is required to achieve errors below 0.1\% in
the gravity gradient components.

The error curve for the computation grid at 260 km height shows that
the error is smaller at higher altitudes.

Computation at higher altitudes would not change the chosen values of the
distance-size ratio for the potential and $g_z$.

The computation height changes the value of the distance-size ratio required
for the gravity gradient components.

We computed the error curves for $g_{zz}$ only at different heights
(Figure~\ref{fig:gzz_error_height}).

The distance-size ratio can be reduced to 4 if the computation is done at 50 km
height.

One can use a distance-size ratio of 2.5 for computations at 260 km height.

These reductions in the distance-size ratio can reduce the computation time
required to achieve the desired.

The software \textit{Tesseroids} includes default values of the distance-size
ratio.

We used the values required to achieve 0.1\% accuracy: 1 for the potential,
1.5 for the gravitational attraction, and 8 for the gravity gradient
components.

The value of $D=8$ for the gravity gradients is conservative.

We chose it to ensure the accuracy of the computation in most cases.

The experienced user of the software can set smaller values using command-line
options.

The smaller values result in less computation time.
